### 2016-04-17 ch03-lrm/LPMSmooth_ML_EG_JS_JL.tex
LPM-trunc finished

The estimated impulse response function  $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$ is obtained explicitly via the \gls{IDFT} of the \gls{LPM}-estimate $\hat G_{\LPM}(\omega_k)$ of the \gls{FRF} in accordance with equation \eqref{eq:defDFT}, \emph{viz}:
\begin{align}\label{eq:impRespiDFT}
\hat g_{{\LPM}\setminus \mathrm{DC
}}(t) &= \frac{1}{\sqrt{N}}\sum_{k=1}^{N-1}\hat G_{\LPM}(\omega_k)e^{\frac{j2\pi kt}{N}}
\end{align}
where the estimated \gls{FRF} in the frequency band between  the Nyquist and the sample frequencies is obtained as follows:
\begin{align}
\hat G_{\LPM}(\omega_{N-k}) = \overline{\hat G_{\LPM}(\omega_k)},\ \text{for}\ k=1,\dots,N/2
\end{align}
(with $\conj{\hat G_{\LPM}}$ the complex conjugate of $\hat G_{\LPM}$) to ensure $\hat g_{\LPM}(t)$ to be real.

Smoothing of the estimated \gls{FRF} requires the correct estimate of the impulse response. 
Unfortunately, the \gls{LPM} presented in \TODO{\secref{sec:nparam:trunc:theoryLPMandWindowing}} and in \citep{Schoukens2009LPM} does not estimate the \gls{FRF}  at the frequency $\omega_0$ (i.e. \gls{DC} value of the \gls{FRF}), hence the subscript ``$\setminus\mathrm{DC}$'' in equation \eqref{eq:impRespiDFT}. 
Consequently, the mean value of the corresponding estimated impulse response given in equation \eqref{eq:impRespiDFT} is not correct. 
This limitation is lifted by developing a simple estimator of the mean value, presented below.

@@

\begin{enumerate}
\item Compute the mean value $m_{g10}$ of the last 10\% of the impulse response $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$, estimated by the \gls{LPM}, as %$\hat g_{\LPM}(t)$ as
\begin{align}
m_{g10} = \frac{1}{\lceil0.1N - 1\rceil}\sum_{t=\lfloor0.9N\rfloor}^{N-1}\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)
\end{align}

\item Next, subtract the computed mean value $m_{g10}$ from $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$ to obtain the improved impulse response $\hat g_{\LPM}(t)$, \emph{viz}:

\begin{align}
\hat g_{\LPM}(t) = \hat g_{{\LPM}\setminus \mathrm{DC}}(t) - m_{g10},\ \text{for}\ t=0,1,\dots,N-1
\end{align}
\end{enumerate}

@@

For the $F$-test based method, the obtained $\truncTime$ values only take a discrete set of values due to the limitation of the segment length.
While most trials result in $\truncTime \in \Set{90, 80, 70}$, a few outliers can be observed.
This can be attributed to the high power of the $F$-tests that are performed.

@@

\begin{guideline}[Check and reinforce assumptions in the data.]
As illustrated by the presented smoothing operation, it pays off to check typical assumptions of the system (e.g. smoothness of the \gls{FRF}, finite support of the impulse response) and even artificially reinforce such assumptions to improve the model quality.
\end{guideline}

\paragraph{Alternative system}
%% BEGIN BOXCAR
\TODO{write about boxcar average, add reference}
\begin{align}
  y(t) = \sum_{i=1}^{\Nbc} \Nbc^{-1} u(t-i)
\end{align}

\begin{figure}
   \centering
        \setlength\figurewidth{0.68\columnwidth}
        \setlength\figureheight{0.68\figurewidth}
        \input{\thisDir/figs/expFitBoxAverager.tikz}
         \caption{Exponential fit of the impulse response of a boxcar averager.}
         \label{fig:nparam:trunc:boxcar:impresp}
\end{figure}

\begin{figure}
   \centering
        \setlength\figurewidth{0.68\columnwidth}
        \setlength\figureheight{0.68\figurewidth}
        \input{\thisDir/figs/pdfs-rmse-boxcar.tikz}
         \caption[RMS error of the FRF versus truncation time (boxcar)]{RMSE of a boxcar averager as a function of the truncation time $\truncTime$.
         Empirical histograms of $\truncTime$ as determined using both the $F$-test and exponential fitting are shown.
         For this system, the $F$-test based method is clearly superior.
         The exponential fitting method is not effective in most cases.}
         \label{fig:nparam:trunc:boxcar:trunctime}
\end{figure}
%% END BOXCAR

@@

The section introduced a novel time domain method to smooth the \gls{LPM}-estimate of an \gls{FRF}. 
It consisted of, after obtaining the \gls{FRF} from the \gls{LPM}, computing the associated estimated impulse response via the \gls{IDFT}.
Then, it was determined statistically at which time index the impulse response had decayed below the noise floor, yielding a point beyond which the response may be set to zero.

>>>>

The estimated impulse response function  $\hat g_{{\LPM}\setminus \mathrm{DC}}(t)$ is obtained explicitly via the \gls{IDFT} of the \gls{LPM}-estimate $\hat G_{\LPM}(\omega_k)$ of the \gls{FRF} in accordance with equation \eqref{eq:defDFT}, \emph{viz}:
\begin{align}\label{eq:impRespiDFT}
\hat g_{{\LPM}\setminus \mathrm{DC
}}(t) &= \frac{1}{\sqrt{N}}\sum_{k=1}^{N-1}\hat G_{\LPM}(\omega_k)e^{\frac{j2\pi kt}{N}}
\end{align}
where the estimated \gls{FRF} in the frequency band between  the Nyquist and the sample frequencies is obtained as follows:
\begin{align}
\hat G_{\LPM}(\omega_{N-k}) = \overline{\hat G_{\LPM}(\omega_k)},\ \text{for}\ k=1,\dots,N/2
\end{align}
(with $\conj{\hat G_{\LPM}}$ the complex conjugate of $\hat G_{\LPM}$) to ensure $\hat g_{\LPM}(t)$ to be real.

Smoothing of the estimated \gls{FRF} requires the correct estimate of the impulse response. 
Unfortunately, the \gls{LPM} presented in \TODO{\secref{sec:nparam:trunc:theoryLPMandWindowing}} and in \citep{Schoukens2009LPM} does not estimate the \gls{DC} value of the \gls{FRF}, hence the subscript ``$\setminus\mathrm{DC}$'' in equation \eqref{eq:impRespiDFT}. 
Consequently, the mean value of the corresponding estimated impulse response given in equation \eqref{eq:impRespiDFT} is not correct. 
This limitation is lifted by developing a simple estimator of the mean value, presented below.

@@

\begin{enumerate}
\item Compute the mean value $m_{g10}$ of the last 10\% of the impulse response $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$, estimated by the \gls{LPM}, as %$\hat g_{\LPM}(t)$ as
\begin{align}
   m_{g10} 
   = 
   \frac{1}{\lceil0.1N - 1\rceil}
   \sum_{t=\lfloor0.9N\rfloor}^{N-1}
      \hat g_{{\LPM}\setminus \mathrm{DC}}(t)
\end{align}

\item Next, subtract the computed mean value $m_{g10}$ from $\hat g_{{\LPM}\setminus \mathrm{DC
}}(t)$ to obtain the improved impulse response $\hat g_{\LPM}(t)$, \emph{viz}:
\begin{align}
\hat g_{\LPM}(t) = \hat g_{{\LPM}\setminus \mathrm{DC}}(t) - m_{g10},\ \text{for}\ t=0,1,\dots,N-1
\end{align}
\end{enumerate}

@@

For the $F$-test based method, the obtained $\truncTime$ values only take a discrete set of values due to the limitation of the segment length.
While most trials result in $\truncTime \in \Set{70, 80, 90}$, a few outliers with much higher truncation times can be observed.
These outliers can be attributed to the high power of the $F$-tests.

@@

\paragraph{Interaction of the truncation method and the system properties}
As seen in the previous simulation example, the exponential fitting method is well-suited to determine $\truncTime$ when the underlying system exhibits an exponential decay.
However, if the system under test is not part of this class of systems, the exponential fitting method may provide sub-optimal results or even be ineffective.

Consider instead that we are measuring a boxcar averager, i.e. a moving-average filter.
Such a system is governed by the difference equation:
\begin{align}
  y(t) = \sum_{i=0}^{\Nbc-1} \Nbc^{-1} u(t-i) 
  \text{,}
\end{align}
where $\Nbc = 88$.
Since this system is a trivial \gls{FIR}, it is easy to see that $\truncTime = \Nbc$ provides the optimal truncation point.

We simulate the boxcar averager for a white noise input $u(t)$ and add white noise with standard deviation $\sigma_e \approx 0.1$ to the output $y(t)$ for $2\,048$ samples from which the first $1\,024$ samples are discarded to randomize the state of the system under test.
A single realization of the impulse response function is shown in \figref{fig:nparam:trunc:boxcar:impresp} together with the $\truncTime$ and other relevant quantities of the exponential fitting method.
It is easy to observe that the exponential fitting provides sub-optimal results since it truncates well past the end of the finite impulse response.

\begin{figure}
   \centering
        \setlength\figurewidth{0.68\columnwidth}
        \setlength\figureheight{0.68\figurewidth}
        \input{\thisDir/figs/expFitBoxAverager.tikz}
         \caption{Exponential fit of the impulse response of a boxcar averager. $\truncTime \gg \Nbc$ shows that a sub-optimal truncation point is found.}
         \label{fig:nparam:trunc:boxcar:impresp}
\end{figure}

By repeating the experiment above $\nMC = 1\,000$ times as a Monte Carlo simulation where both the input signal and disturbing noise are varied, we gain insight into the behavior of both truncation methods.
The histograms of the observed values of $\truncTime$ for both methods is shown in \figref{fig:nparam:trunc:boxcar:trunctime} together with the \gls{RMS} error as a function of $\truncTime$.
It can be seen that the $F$-test based method mostly selects the optimal truncation point.
On the other hand, the exponential fitting method is largely ineffective and doesn't truncate the response in most of the cases.

\begin{figure}
   \centering
        \setlength\figurewidth{0.68\columnwidth}
        \setlength\figureheight{0.68\figurewidth}
        \input{\thisDir/figs/pdfs-rmse-boxcar.tikz}
         \caption[RMS error of the FRF versus truncation time (boxcar)]{RMSE of a boxcar averager as a function of the truncation time $\truncTime$.
         Empirical histograms of $\truncTime$ as determined using both the $F$-test and exponential fitting are shown.
         For this system, the $F$-test based method is clearly superior.
         The exponential fitting method is not effective in most cases.}
         \label{fig:nparam:trunc:boxcar:trunctime}
\end{figure}

It can be seen that both the $F$-test method and the exponential fitting method each have their own strengths and weaknesses.
Particularly, the $F$-test method can only select $\truncTime$ from a discrete set of values, but works well for specifically designed \gls{FIR} filters where there is a clear jump in the amplitude of the impulse response.
On the other hand, the exponential fitting method relies on the exponential decay of the impulse response to obtain sensible values of $\truncTime$ and is hence a better match for `physical' systems.
One should remember, however, that for the studied examples, both approaches only very seldomly select a $\truncTime$ that increases the \gls{RMS} error on the impulse response compared to not truncating the response.
As such, a use has little to lose by trying out both methods: the model quality is very unlikely to decrease and the additional computational effort is minimal.

\begin{guideline}[Check and reinforce assumptions in the data.]
As illustrated by the presented smoothing operation, it pays off to check typical assumptions of the system (e.g. smoothness of the \gls{FRF}, finite support of the impulse response) and even reinforce such assumptions to improve the model quality.
For the truncation approaches, even if the assumptions do not hold, the methods are constructed such that the unavoidable model error is unlikely to increase.
\end{guideline}

@@

This section introduced a time domain method to smooth the \gls{LPM}-estimate of an \gls{FRF}. 
It consisted of, after obtaining the \gls{FRF} from the \gls{LPM}, computing the associated estimated impulse response via the \gls{IDFT}.
Then, it was determined statistically at which time index the impulse response had decayed below the noise floor, yielding a point beyond which the response may be set to zero.
